\begin{refsection}
\chapter{Summary of the dissertation}
\label{ch1}
The Great Recession following the financial crisis of 2007-2008 reinvigorated the interest in the research of financial development and its impact on the real economy. Although it was a dominant view at the time that more finance is good for economic growth and equalizes opportunities, the crisis spurred questions about the non-linearities in finance effects and existence of a \emph{healthy} threshold or the impact being conditional on the quality of institutions. New regulatory waves followed focusing primarily on the stability of the individual financial intermediaries, but also the systemic risk in general. 

The measurement of finance and financial development also came in question. Most of the researchers in the field relied on the size (depth) measurement of finance. These are, however, imperfect proxies of the functions ascribed to finance in the theoretical models \parencite{Levine2005}. When I was about to begin with my dissertation research, more detailed and sufficiently dense data on financial development were published by the World Bank. The information on the stability, efficiency, and access of financial intermediaries was not perfect, but it nevertheless indicated better alternatives to size in accurately capturing the channels through which finance affects the real economic phenomena. Together with the carefully adopted methodological approach in \ac{BMA}, which allows for a comparison of relative importance of different financial proxies, it suggested a promising research path I decided to take. Given the relevance of the question at the time, economic growth was an easy first choice. The evolution of the debate on inequality in the aftermath of financial crisis meant that the topics of wealth and income inequality naturally followed.

This thesis is a composition of three papers related to financial development and its consequences. When I mention financial development in this thesis, I refer to the ability of financial contracts, markets, and intermediaries to facilitate the screening of investment opportunities; the monitoring of investments; and the pooling, trading, and management of risk \parencite{demirgucc2009finance}. In selected chapters that follow, I often reference to the authorship as ``we'' which reflects that some of the papers in this dissertation I wrote in collaboration with my supervisor professor Roman Horv\'{a}th and professor Iftekhar Hasan from Fordham University. I continue with the overviews of individual dissertation papers.

In \emph{Chapter 2 - What Type of Finance Matters for Growth? Bayesian Model Averaging Evidence}, we examine the effect of finance on long-term economic growth. We consider the size proxies jointly with indicators that assess the stability and efficiency of financial markets. In the paper, we address the inconclusive finance-growth nexus literature. While some claim financial development has positive effect on economic growth \parencite{AtjeJovanovich1993,KingLevine1993a,RajanZingales1998}, others hold that financial sector removes scarce resources from the economy \parencite{bolton2016cream,Tobin1984,axelson2015wall} and underpins greater exposure and vulnerability to crises, severely burdening the real sector in during periods of instability \parencite{Minsky1991,Stiglitz2000}. More recent papers also point towards decreasing returns to financial development and finance having negative consequences for growth when above certain threshold \parencite{Arcandetal2012,LawSingh2014,RousseauWachtel2011}. 

We depart from the literature in two main features. First, we apply \ac{BMA} to solve the model uncertainty problem in growth regressions. The variety of theories of economic growth suggests a large number of determinants and results in considerable uncertainty about the ``true'' growth model. Using the \ac{BMA}, we can evaluate numerous regressors potentially relevant for economic growth and estimate their \ac{PIP}, the probability that they are relevant in explaining the dependent variable, additionally to the weighted mean and variance of the respective coefficients. \ac{BMA} essentially estimates varying combinations of explanatory variables and weights the coefficients using model fit. The methodology is solidly rooted in the statistical theory \parencite{Rafteryetal1997,Koopetal2007} and indirectly also helps us to tackle the potential of omitted variables bias, which is typically abstracted from in the empirical work on finance and growth.

Second, we augment previous research by examining several financial development indicators simultaneously to account for multidimensionality of financial systems. By jointly examining whether depth, stability, or efficiency is relevant for long-term growth, we re-examine and unify previous literature. The theoretical functions of finance are difficult to operationalize in the empirical research \parencite{Valickovaetal2014} and there is no consensus on the measurement of financial of financial development \parencite{KingLevine1993a}. The research dominantly uses depth of financial markets (credit / GDP ratio or stock market capitalization / GDP) as a measure of financial development. Employing the \ac{GFDD} and the indicators provided therein, we can approximate the function of financial system in much more detail. We can discriminate between banking and stock markets as well as evaluate the relative importance of depth versus the alternative proxies of efficiency, stability, and access to finance. We may also reflect the claims that excessive financial development and financial instability are harmful for growth. Even thought he data coverage is still somewhat limited, we contribute to the literature by considering these additional dimension of the financial sector in our regression analysis to provide a more complete picture of finance-growth nexus. We complement the data on financial development by the long-term growth dataset of \textcite{Fernandezetal2001}, which provides a rich set of possible explanatory variables capturing various economic, political, geographical and institutional factors.

We find that bank efficiency is the only indicator of financial development which is robustly related to economic growth and consistently shows very high \ac{PIP}. This result is consistent with the theoretical predictions sketched out by \textcite{Pagano1993} who shows how the increased efficiency of financial intermediaries affects the channel between savings and investment and therefore leads to higher real growth. On the other hand, the relevance of the traditionally employed variables, such as credit to the private sector or stock market capitalization, is weaker. Additionally, we find no evidence for a non-linear effect of financial development. We subject our results to further robustness checks by focusing on different sample period, employing alternative priors, and elementary techniques to address endogeneity with no substantial effect on our conclusions. The policy relevance of the results is in highlighting the essential importance of measuring financial development to precisely describe its consequences. The regulatory changes in financial industry should appreciate the relevance of the financial intermediaries for long-term growth.

\emph{Chapter 3 - Finance and Wealth Inequality} extends the idea of distinct features of financial systems to the distribution of wealth. Wealth inequality markedly varies across countries \parencite{daviesetal2017} and the interest of the paper is to uncover the drivers of these differences. Is it different degrees of redistribution, financial development, globalization, technological progress, education, economic development, or something else? Although important progress has been made in advancing the measurement of wealth inequality \parencite{daviesetal2017,SaezZucman2016}, there is lack of systematic evidence about the determinants of wealth inequality across countries.

The theoretical predictions of the wealth inequality drivers vary. Much discussed $r > g$ concept presented by \textcite{piketty2014} suggests there is a natural tendency towards increasing wealth inequality unless exogenously amended by redistribution or wars. The framework is criticized on many fronts though \parencite{mankiw2015yes,blume2015capital,king2017literature} and the corner stone remains with distinct applications of dynamic quantitative models. The models\footnote{\textcite{DENARDI2017280} provide thorough overview of the model implementations.} critically rely on the saving motives of the individuals and this leads us to hypothesis of the financial development being strongly relevant for wealth distribution. Another prediction about the financial system and wealth inequality arises from \textcite{pastor2016income} in which inequality is driven, among other things, by the ability of entrepreneurs to diversify away their idiosyncratic risk. The empirical evidence is scarce as the research papers on inequality mostly turn to the distribution of income due to better data availability. There is evidence, however, that wealth is much more unevenly distributed than income \textcite{zucman2019,oecd2013crisis} and income distribution is mostly used as an approximation of wealth distribution, while the latter would be more fitting given the theory \parencite{bagchi2015does}.

The lack of encompassing theoretical framework informs our methodological framework similarly as in the paper on finance and growth. We rely on \ac{BMA} in estimations to identify relevant determinants of wealth distribution. Moreover, we extend the analysis to address potential endogeneity more rigorously using the \ac{IVBMA}. \ac{IVBMA} essentially resembles the two-stage frequentist methods, but accounts for model uncertainty in both stages. We further refresh and expand the set of regressors by constructing our original database although conceptually, we similarly select the variables capturing economic, financial, institutional, regulatory, and political features of considered countries. We prefer the freedom about the choice of regressors over the comparability of our results with existing research as the paper is a pioneering work in this field. We also importantly update the indicators of financial development we employ. Rather than relying on single indicators capturing different dimensions of financial systems, we use complex indicators constructed from the \ac{GFDD} that describe the characteristics of financial systems by extracting information from multiple indicators in each dimension through principle component analysis.

We find that the set of key determinants is small and financial development exerts a complex effect on wealth inequality. Whereas countries with deeper financial systems (large financial markets and financial institutions) exhibit greater wealth inequality, more efficient intermediation and access to finance is associated with associated with less wealth inequality. Our results thus support the idea that sound financial systems may contribute to lower wealth inequality. Alongside financial development, we discover that education, redistribution, globalization, and political instability affect wealth distributions within countries. Better educated societies and more higher redistribution of income support more egalitarian distribution of wealth while globalization and political instability increase wealth inequality. The conclusions offer clear policy alternatives of countermeasures to increasingly unequal distributions of wealth in inclusive and efficient financial systems alongside better education.

\emph{Chapter 4 - Finance and Inequality - panel BMA approach} is the last follow-up in the series of papers on financial development. The theoretical predictions and ambiguity resembles the ones on wealth inequality and finance. In contrast with wealth, income inequality and the relation to finance has been explored much more frequently, but nevertheless with conflicting outcomes. A key divide appears between financial development on extensive and intensive margin. On the extensive margin, it might lead to more equal opportunities and outcomes as access to credit by previously disadvantaged groups allows human capital accumulation \parencite{braunetal2019,galormoav2004} and formation of new firms \parencite{banerjeenewman1990,evans1989estimated}. On the contrary, intensive margin of financial development might inordinately benefit the rich incumbents who leverage financial services for their further benefit or to protect their existing rents \parencite{GreenwoodJovanovic1990,perotti2007investor}.

The paper re-examines the literature on finance and inequality by applying panel \ac{BMA} techniques, once more identify the key determinants of income distribution within countries. I contribute to existing research by showing that: 1) finance has significant role in shaping the distribution of income, 2) the complexity lies in the characteristics of financial systems, and 3) the effect varies across different parts of income distribution. 

Reflecting the conclusions of the preceding chapter, efficiency and access to financial institutions appears to have the inequality reducing role. The depth of financial system seemingly does not influence the overall measure of income distribution (Gini index), but when we focus on the top income shares, we can associate size of the financial markets and institutions with more concentrated distribution of income. Additionally to the results on financial development, we also provide evidence on other popular hypotheses about increasing income inequality exploring the education \parencite{goldin2009race}, globalization \parencite{Jaumotte2013}, or technological progress \parencite{dabla2015causes}.

% Relying on the global sample of countries. We identify many potentially relevant regressors hypothesized in the literature as potential drivers of wealth distribution. 

% Financial development debate following the crisis\dots

% \ac{BMA} useful in examining uncertain relationships\dots + what is it, well grounded in theory

% Roles of finance\dots + multidimensionality

\newpage
\printbibliography[heading=subbibliography]
\addcontentsline{toc}{section}{References}
\end{refsection}